Manuscript_Github_ancestral_pir_210801

### 2021-09-21 AncestralPir/AnalysisCode/Manuscript_Github_AncestralPir.Rmd
update renv snapshot and update ancestral pir code to source functions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode")
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

@@

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r tpm_function}
tpm = function(count_data, lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      print('Warning:Count data and length gene ids DO NOT MATCH')
    }
  } else {
    print('Count data and/or length table supplied do not have an id column, 
          make sure they are ordered correctly')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = TRUE))){
    gene_names <- count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
    counts_tpm <- count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
  }else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = TRUE))){
    lengths.table = lengths[,grep(colnames(lengths), pattern = 'length', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  #Making lengths.table a matrix sometimes throws up problems
  lengths.table = as.matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  colnames(counts_tpm) <- 'tpm'
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), 
                                            pattern = 'Geneid', 
                                            ignore.case = T), 
                                      drop = FALSE],
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

@@

coat_transcript_lengths_path <- list.files('../../genomes_genes_annotations/', 
                      pattern = 'PlasmoDBv48_Pcoatneyi',
                      full.names = TRUE)
coat_transcript_lengths <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(c(Geneid, `Transcript Length`))

@@

(coat_tpm$ancestral_ortholog %>% sum)
coat_anc_v_nonanc <- coat_tpm %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('tpm')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
                   standev = sd(value))
coat_data <- data.frame(Geneid = coat_tpm$Geneid[which(coat_tpm$ancestral_ortholog == 1)],
                        ratio = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)]/
                          coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)])
```

@@

#This gives 5074.5 so need to make sure that not having an integer count messes up things but it shouldn't
yoel_lengths <- yoel_lengths %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(length = median(length))
nrow(yoel_lengths)
nrow(yoel_counts)

yoel_tpm <- tpm(count_data = yoel_counts, lengths = yoel_lengths)
yoel_tpm <- yoel_tpm %>% 
  mutate(Description = yoel_counts_full$Description[match(Geneid, 
                                                          yoel_counts_full$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

@@

```{r sourcing_functions}
source('../../rfunctions/uniquify_fun.R')
source('../../rfunctions/ordering_fun.R')
source('../../rfunctions/tpm_fun.R')
source('../../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
source('../../rfunctions/ecdf_percentile_fun.R')
```

@@

coat_transcript_lengths_path <- list.files('../../genomes_genes_annotations/', 
                      pattern = 'PlasmoDBv48_Pcoatneyi',
                      full.names = TRUE)
coat_transcript_lengths <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  rename(Transcript.Length = `Transcript Length`) %>% 
  select(c(Geneid, Transcript.Length ))

@@

(coat_tpm$ancestral_ortholog %>% sum)
coat_anc_v_nonanc <- coat_tpm %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('WB')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
                   standev = sd(value))
coat_data <- data.frame(Geneid = coat_tpm$Geneid[which(coat_tpm$ancestral_ortholog == 1)],
                        ratio = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)]/
                          coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)])
```

@@

#This gives 5074.5 so need to make sure that not having an integer count messes up things but it shouldn't
yoel_lengths <- yoel_lengths %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(Transcript.Length  = median(length))
nrow(yoel_lengths)
nrow(yoel_counts)

yoel_tpm <- tpm(count_data = yoel_counts, lengths = yoel_lengths)
yoel_tpm <- yoel_tpm %>% 
  mutate(Description = yoel_counts_full$Description[match(Geneid, 
                                                          yoel_counts_full$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0)) %>% 
  rename(tpm = Counts)
